---
title: "Exercise 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls()) # clear memory, step 1
gc() # clear memory, step 2
```

## Setting up data and environment

We first need to do a few things before we can manipulate the data. 

```{r data}
# set path for R to find our data
data_path <- "~/Dropbox/McGill/teaching/2021-2022/2022_summer/ORGB672/data/"
setwd(data_path)

# load the necessary packages ("extensions")
library(tidyverse)
library(lubridate) # to work with dates
library(skimr) # for summaries of data
library(arrow) # to be able to load data in the .parquet format

# read application data
app_data_sample <- read_parquet(paste0(data_path,"app_data_sample.parquet"))
```

## Let's find the first and the last observed date for each examiner

We'll first get examiner IDs and application dates in a separate dataframe (table), for ease of manipulation. We'll keep examiner ID (the field `examiner_id`), and earliest and latest dates for each application (`filing_date` and `appl_status_date` respectively).

```{r}
examiner_dates <- app_data_sample %>% 
  select(examiner_id, filing_date, appl_status_date) 

examiner_dates
```

The dates look inconsistent in terms of formatting. Let's make them consistent. We'll create new variables `start_date` and `end_date`.

```{r}
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))
```

Let's now identify the earliest and the latest date for each examiner and calculate the difference in days, which is their tenure in the organization.

```{r}
examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure = interval(earliest_date, latest_date) %/% days(1)
    )

examiner_dates

skim(examiner_dates)
```

## Find examiner gender
##### >> add examiner gender -----
# Using a modified example from https://cran.r-project.org/web/packages/gender/vignettes/predicting-gender.html
library(gender)
#install_genderdata_package() # only run the first time
examiner_gender <- examiner_gs %>% 
  mutate(
    name = examiner_name,
    ff_name = str_extract(examiner_name,"(?<=,\\s)\\w+"), # extract first first name
  ) %>% 
  distinct(ff_name) %>% 
  do(results = gender(.$ff_name, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    ff_name = name,
    gender,
    proportion_female
  )
# joining gender back to the dataset
examiner_gs <- examiner_gs %>% 
  mutate(
    ff_name = str_extract(examiner_name,"(?<=,\\s)\\w+"), # extract first first name
  ) %>% 
  left_join(examiner_gender, by = "ff_name")
# cleaning up
rm(examiner_gender)
gc()
