---
title: "Exercise 4 quick start"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidygraph)
library(ggraph)
```

## Illustration of friendship network

We'll use `tidygraph` package to create and plot the friendship network from the slides. The nodes are Abby (A), Becca (B), Chloe (C) and Deb (D). Abby is only friends with Becca; Becca is friends with Abby, Chloe and Deb; and Chloe and Deb and friends with Becca and each other (but not Abby).

We'll be relying on these tutorials:

-   <http://users.dimi.uniud.it/~massimo.franceschet/ns/syllabus/make/tidygraph/tidygraph.html>
-   <https://www.data-imaginist.com/2017/introducing-tidygraph/>
-   <https://www.data-imaginist.com/2018/tidygraph-1-1-a-tidy-hope/>

First, we define the network by manually entering all the nodes and the connections between them. We'll define `nodes` table that has two columns: `id` and `name`. We will then define the connections among them as an edgelist, where each element in column `from` corresponds to a friend on one end of the relationship and each element in column `to` corresponds to the person on the other end of this friendship tie.

```{r define-network}
# define nodes
node_names <- tibble(
  id   = c(1,2,3,4),
  name = c("Abby","Becca","Chloe","Deb")
)
node_names

# define connections (have to correspond to ties 1-2, 2-3, 2-4, 3-4)
# for each element in `from` there is a corresponding element in `to`
edge_list <- tibble(
  from = c(1,2,2,3),
  to   = c(2,3,4,4)
)
edge_list
```

We can now combine these tables into a "graph" object that holds all of our network data.

```{r define-graph}
# combine this information into a network graph object
friendship_graph <- tbl_graph(nodes = node_names, edges = edge_list, directed = FALSE)
friendship_graph
```

We can now plot this network using `ggraph` package.

```{r plot-graph}
friendship_graph %>% 
    ggraph(layout = 'kk') + 
    geom_edge_link() + 
    geom_node_point(size = 8, colour = 'gray') +
    geom_node_text(aes(label = name), colour = 'steelblue', vjust = 0.4) + 
    ggtitle('Friendship network') + 
    theme_graph()
```

We can use many of the functions in package `tidy_graph` to calculate things we want to know about this network. For example, we may want to know the centrality of each node in the network.

```{r}
friendship_graph <- friendship_graph %>% 
  activate(nodes) %>% # we need to state we'll be adding to nodes, not edges
  mutate(d_centrality = centrality_degree()) %>%  # adding measure of degree centrality
  mutate(b_centrality = centrality_betweenness()) # adding betweenness centrality

friendship_graph
```

Now let's plot this with degree centrality determining the size of the nodes and betweenness determining its color.

```{r plot-centrality}
friendship_graph %>% 
  ggraph(layout = 'kk') + 
  geom_edge_link() + 
  geom_node_point(aes(size = d_centrality, colour = b_centrality)) + 
  scale_color_continuous(guide = 'legend') +
  geom_node_text(aes(label = name), colour = 'red', vjust = 1.6) + 
  ggtitle('Friendship network') + 
  theme_graph()
```
